Trends from 30-year observations of downward solar irradiance in Thessaloniki, Greece

### 2023-12-19 MDPI_submission.Rmd
Commit 2023-12-19 19:11

## numeric precision of table
LT$slope           <- signif(LT$slope,               2)
LT$slope.p         <- signif(LT$slope.p,             3)
LT$Rsqrd           <- signif(LT$Rsqrd,               2)
LT$Tmod_Estimate   <- signif(LT$Tmod_Estimate,       2)
LT$cor.estimate    <- signif(LT$cor.estimate,        2)
LT$Tmod_Std..Error <- signif(LT$Tmod_Std..Error * 2, 2)
LT$ChangeWpY       <- signif(LT$ChangeWpY,           2)
LT$slope <- format(LT$slope, 3, drop0trailing = T)

@@

```{r trendALL, echo=F, out.width='.75\\linewidth', fig.cap=paste0("Anomalies (\\%) of the daily all-sky SDR from the climatological mean for the period ", year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)), ". The black line is the long term linear trend.")}
knitr::include_graphics("./images/LongtermTrends-2.png")
```

@@

We have analyzed a 30 year dataset of SRD measurements in Thessaloniki Greece 
(`r paste(year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)))`) aiming to identify the long term variability of
solar irradiance under different sky conditions.
Under all-sky conditions 
there is a positive trend in SDR of $`r alltrend`\,\%/y$ (brightening). 
A previous study [@Bais2013] for the period 1993 -- 2011 reported also a positive trend, of $0.33\,\%/y$.
The slight increase of this trend indicates that the brightening of SDR continues and is probably caused by continuing decreases in aerosol optical depth and the optical thickness of clouds over the area.
A smaller trend has
been found under clear-sky conditions ($`r cleartrend`\,\%/y$) which supports the
notion that part of the brightening is caused by decreasing aerosols.
@Siomos2020 have shown that aerosol optical depth over
Thessaloniki is decreasing constantly at least up to 2018.
Unfortunately, for this study aerosol data for the entire period were
not available in order to quantify their effect on SDR. The attenuation
of SDR by aerosols over Europe has been proposed as major factor by @Wild2021.
The brightening effect on SDR under cloudy-sky conditions
($`r cloudtrend`\,\%/y$), suggests that cloud optical thickness is also decreasing
during this period.
As long term data of cloud optical thickness are
also not available for the region, we cannot draw quantitative
conclusions.

>>>>

## numeric precision of table
LT$slope           <- signif(LT$slope,               2)
LT$slope.p         <- signif(LT$slope.p,             3)
LT$Rsqrd           <- signif(LT$Rsqrd,               2)
LT$Tmod_Estimate   <- signif(LT$Tmod_Estimate,       2)
LT$cor.estimate    <- signif(LT$cor.estimate,        2)
LT$Tmod_Std..Error <- signif(LT$Tmod_Std..Error * 2, 2)
LT$ChangeWpY       <- signif(LT$ChangeWpY,           3)
LT$slope <- format(LT$slope, 3, drop0trailing = T)

@@

```{r trendALL, echo=F, out.width='.75\\linewidth', fig.cap=paste0("Anomalies (\\%) of the daily all-sky SDR from the climatological mean for the period ", year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)), ". The black line is the long term linear trend.")}
knitr::include_graphics("./images/LongtermTrends-1.png")
```

@@

We have analyzed a 30 year dataset of SRD measurements in Thessaloniki Greece 
(`r paste(year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)))`) aiming to identify the long term variability of
solar irradiance under different sky conditions.
Under all-sky conditions 
there is a positive trend in SDR of $`r alltrend`\,\%/y$ (brightening). 
A previous study [@Bais2013] for the period 1993 -- 2011 reported also a positive trend, of $0.33\,\%/y$.
The slight increase of this trend indicates that the brightening of SDR continues and is probably caused by continuing decreases in aerosol optical depth and the optical thickness of clouds over the area.
A smaller trend has
been found under clear-sky conditions ($`r cleartrend`\,\%/y$) which supports the
notion that part of the brightening is caused by decreasing aerosols.
@Siomos2020 have shown that aerosol optical depth over
Thessaloniki is decreasing constantly at least up to 2018.
Unfortunately, for this study aerosol data for the entire period were
not available in order to quantify their effect on SDR. The attenuation
of SDR by aerosols over Europe has been proposed as major factor by @Wild2021.
The brightening effect on SDR under cloudy-sky conditions
($`r cloudtrend`\,\%/y$), suggests that cloud optical thickness is also decreasing
during this period.
As long term data of cloud optical thickness are
also not available for the region, we cannot draw quantitative
conclusions.
@Lozano2023 reported a Cloud Radiative Forcing (CRF) trend of $1.22\,W/m^2/y$ for the SDR for the city of Granada Spain ($37.16^\circ\,$N, $3.61^\circ\,$E, $680\,$m\ a.s.l.). 
 <!-- (37.16◦ N, 3.61◦ W, 680 m a.s.l.), 37° 9' 35.9994" N  3° 36' 35.9994" W-->
We found a comparable result, for the same parameter, using the long term trend of all- and clear-skies, as follows $CRF = 1.46 - 0.501 = `r 1.46 - 0.501`\,W/m^2/y$.

### 2023-12-19 MDPI_submission.tex
Commit 2023-12-19 19:11

\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
Sky conditions & Trend [\%/year] & Trend S.E. ($2\sigma$) & Pearson correl. & Trend [W/m\textsuperscript{2}/year] & Days\\
\midrule
All skies & 0.380 & 0.120 & 0.091 & 1.5 & 10251\\
Clear skies & 0.097 & 0.033 & 0.140 & 0.5 & 2684\\
Cloudy skies & 0.410 & 0.180 & 0.081 & 1.2 & 4937\\
\bottomrule
\end{tabu}
\end{table}

@@

{\centering \includegraphics[width=.75\linewidth]{./images/LongtermTrends-2}

@@

We have analyzed a 30 year dataset of SRD measurements in Thessaloniki
Greece (1993 -- 2023) aiming to identify the long term variability of
solar irradiance under different sky conditions. Under all-sky
conditions there is a positive trend in SDR of \(0.38\,\%/y\)
(brightening). A previous study \citep{Bais2013} for the period 1993 --
2011 reported also a positive trend, of \(0.33\,\%/y\). The slight
increase of this trend indicates that the brightening of SDR continues
and is probably caused by continuing decreases in aerosol optical depth
and the optical thickness of clouds over the area. A smaller trend has
been found under clear-sky conditions (\(0.097\,\%/y\)) which supports
the notion that part of the brightening is caused by decreasing
aerosols. \citet{Siomos2020} have shown that aerosol optical depth over
Thessaloniki is decreasing constantly at least up to 2018.
Unfortunately, for this study aerosol data for the entire period were
not available in order to quantify their effect on SDR. The attenuation
of SDR by aerosols over Europe has been proposed as major factor by
\citet{Wild2021}. The brightening effect on SDR under cloudy-sky
conditions (\(0.41\,\%/y\)), suggests that cloud optical thickness is
also decreasing during this period. As long term data of cloud optical
thickness are also not available for the region, we cannot draw
quantitative conclusions.

>>>>

\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
Sky conditions & Trend [\%/year] & Trend S.E. ($2\sigma$) & Pearson correl. & Trend [W/m\textsuperscript{2}/year] & Days\\
\midrule
All skies & 0.380 & 0.120 & 0.091 & 1.460 & 10251\\
Clear skies & 0.097 & 0.033 & 0.140 & 0.501 & 2684\\
Cloudy skies & 0.410 & 0.180 & 0.081 & 1.180 & 4937\\
\bottomrule
\end{tabu}
\end{table}

@@

{\centering \includegraphics[width=.75\linewidth]{./images/LongtermTrends-1}

@@

We have analyzed a 30 year dataset of SRD measurements in Thessaloniki
Greece (1993 -- 2023) aiming to identify the long term variability of
solar irradiance under different sky conditions. Under all-sky
conditions there is a positive trend in SDR of \(0.38\,\%/y\)
(brightening). A previous study \citep{Bais2013} for the period 1993 --
2011 reported also a positive trend, of \(0.33\,\%/y\). The slight
increase of this trend indicates that the brightening of SDR continues
and is probably caused by continuing decreases in aerosol optical depth
and the optical thickness of clouds over the area. A smaller trend has
been found under clear-sky conditions (\(0.097\,\%/y\)) which supports
the notion that part of the brightening is caused by decreasing
aerosols. \citet{Siomos2020} have shown that aerosol optical depth over
Thessaloniki is decreasing constantly at least up to 2018.
Unfortunately, for this study aerosol data for the entire period were
not available in order to quantify their effect on SDR. The attenuation
of SDR by aerosols over Europe has been proposed as major factor by
\citet{Wild2021}. The brightening effect on SDR under cloudy-sky
conditions (\(0.41\,\%/y\)), suggests that cloud optical thickness is
also decreasing during this period. As long term data of cloud optical
thickness are also not available for the region, we cannot draw
quantitative conclusions. \citet{Lozano2023} reported a Cloud Radiative
Forcing (CRF) trend of \(1.22\,W/m^2/y\) for the SDR for the city of
Granada Spain (\(37.16^\circ\,\)N, \(3.61^\circ\,\)E,
\(680\,\)m~a.s.l.). We found a comparable result, for the same
parameter, using the long term trend of all- and clear-skies, as follows
\(CRF = 1.46 - 0.501 = 0.959\,W/m^2/y\).
